\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-12 paper/ms.tex
More writing about acceptance probability

Diffusive Nested Sampling \citep[DNS][]{dnest} is an alternative to NS for
problems where MCMC is the only viable sampling method. DNS is based on the
Metropolis algorithm. DNS evolves a point in the parameter space, along with
an integer index variable $j$,
to explore the following joint distribution:
\begin{align}
p(\theta, j) &= p(j)p(\theta | j)\\
&= w_j \times
\frac{\pi(\theta)\mathds{1}\left[L(\theta) > \ell_j\right]}{X(\ell_j)}.
\end{align}
where $\{\ell_j\}$ is a sequence of increasing likelihood thresholds
or {\em levels}, and
$\ell_0 = 0$, and $\{w_j\}$ is the marginal distribution for $j$.
The marginal distribution for $\theta$ is then a {\em mixture} of
constrained priors:
\begin{align}
p(\theta) &=
\pi(\theta)\sum_{j=0}^{j_{\rm max}}
\frac{w_j\mathds{1}\left[L(\theta) > \ell_j\right]}{X(\ell_j)}.
\label{eqn:mixture_of_constrained_priors}
\end{align}

@@

In Nested Sampling, the target distribution is not the posterior but rather
the constrained prior. Hence, the appropriate acceptance probability is
\begin{align}
\alpha &= \min\left[1,
\frac{q(\params'|\params)}{q(\params | \params')}
\frac{\pi(\params')}{\pi(\params)}
\mathds{1}\left(L(\params') > L^*\right)
\right]\\
&= \min\left[1,
e^{\ln(H)}
\mathds{1}\left(L(\params') > L^*\right)
\right]\label{eqn:log_hastings}
\end{align}
where $L^*$ is the threshold value of the likelihood.

>>>>

Diffusive Nested Sampling \citep[DNS][]{dnest} is an alternative to NS for
problems where MCMC is the only viable sampling method. DNS is based on the
Metropolis algorithm. DNS evolves a point in the parameter space, along with
an integer index variable $j$,
to explore the following joint distribution:
\begin{align}
p(\theta, j) &= p(j)p(\theta | j)\\
&= w_j \times
\frac{\pi(\theta)\mathds{1}\left[L(\theta) > \ell_j\right]}{X(\ell_j)}.
\label{eqn:target_distribution}
\end{align}
where the $\{\ell_j\}$ are a sequence of increasing likelihood thresholds
or {\em levels}, and
$\ell_0 = 0$, and $\{w_j\}$ is the marginal distribution for $j$.
The marginal distribution for $\theta$ is then a {\em mixture} of
constrained priors:
\begin{align}
p(\theta) &=
\pi(\theta)\sum_{j=0}^{j_{\rm max}}
\frac{w_j\mathds{1}\left[L(\theta) > \ell_j\right]}{X(\ell_j)}.
\label{eqn:mixture_of_constrained_priors}
\end{align}

@@

In DNS, the target distribution is not the posterior but rather
the joint distribution in Equation~\ref{eqn:target_distribution}.
Moves of $\theta$ are done keeping $j$ fixed, so we only need
to consider the Metropolis acceptance probability for fixed $j$,
i.e. with respect to a single constrained prior like
Equation~\ref{eqn:constrained_prior}.
Hence, the appropriate acceptance probability
for a proposed move from $\theta$ to $\theta'$ is
\begin{align}
\alpha &= \min\left[1,
\frac{q(\params'|\params)}{q(\params | \params')}
\frac{\pi(\params')}{\pi(\params)}
\mathds{1}\left(L(\params') > \ell_j\right)
\right]\\
&= \min\left[1,
e^{\ln(H)}
\mathds{1}\left(L(\params') > \ell_j\right)
\right]\label{eqn:log_hastings}
\end{align}
where $\ell_j$ is the likelihood threshold for level $j$.
